```{r load libraries}
library(dplyr)
library(stringr)
library(openxlsx)
library(here)
library(tidyr)
```

Setting Data folder and file locations

```{r data locations and files}
## raw data and wrangled data folder locations -----
raw_data_loc <- here("data")
scripts_loc <- here("scripts")


## Data file locations for convenience -----
data_dictionary_file <- here(raw_data_loc, "Data_dictionary_with_description_20201105_CORRECTED.xlsx")
raw_survey_data_file <- here(raw_data_loc, "SPEAKI_Survey_Data_deidentified.xlsx")
```


Source Functions

```{r}
source(here(scripts_loc, "read-data-dict.R"))
source(here(scripts_loc, "parse-survey.R"))
       
```


Read in Data dictionary sheet

```{r}
data_dictionary <- read_data_dictionary(data_dictionary_file)
```


Read in CHSA information

```{r}
chsa_tbl <- read_chsa_data_dict(data_dictionary_file)
```


Read in LHA information

```{r}
lha_tbl <- read_lha_data_dict(data_dictionary_file)
```


Read in HA information

```{r}
ha_tbl <- read_ha_data_dict(data_dictionary_file)
```


Read in HSDA information

```{r}
hsda_tbl <- read_hsda_data_dict(data_dictionary_file)
```


Read in CHSA UR Class

```{r}
chsa_ur_class_tbl <- read_chsa_ur_class_data_dict(data_dictionary_file)
```

Read in Raw Survey Data

```{r}
## Reading in a sample of the survey data to sample -----
## 100 rows and all the columns 

raw_survey_data <- read.xlsx(raw_survey_data_file,
                             rows = c(1:100)) %>%
  as_tibble()

```


Join Data dictionary with the raw survey data. This can be run directly
without reading in data dictionary file 

```{r}
joined_survey_data_dictionary_tbl <- join_dictionary_with_survey_data(raw_survey_data,
                                                                      data_dictionary_file)
```


Parsing data dictionary values in the survey data. Currently this has
problems while parsing the whole table. 

```{r}
parsed_survey_data_tbl <- parse_survey_data_dictionary(joined_survey_data_dictionary_tbl)
```
