---
title: "CanvasXpress - Maps"
author: "Dr. Connie Brett"
date: "9/20/2021"
output: 
  html_document: 
    toc: yes
    theme: cerulean
params:
    data_dir:              "../data"
    lha_boundaries_file:  "map_LHA_boundaries.json"
    summary_data_file:    "BC_COVID_19_Survey_downloadable_data.xlsx"
    data_dictionary_file: "Data_dictionary_with_description_20201105_CORRECTED.xlsx"
    
---

```{r setup, include=FALSE}
library(canvasXpress)
library(htmlwidgets)
library(jsonlite)
library(dplyr)
library(glue)
library(openxlsx)
library(tibble)

# for debugging purposes (will not execute when knit)
if (interactive()) {
    try({ setwd('markdown') }, silent = T)
    params <- list(data_dir             = "../data",
                   lha_boundaries_file  = "map_LHA_boundaries.json",
                   summary_data_file    = "BC_COVID_19_Survey_downloadable_data.xlsx",
                   data_dictionary_file = "Data_dictionary_with_description_20201105_CORRECTED.xlsx")
}

if (packageVersion('canvasXpress') <= "1.36.1") {
    message('WARNING: update canvasXpress to 1.36.2+ for map usage')
    message('To get the latest cran candiate release use devtools::install_github("cb4ds/canvasXpress@cran")')
}

```



## CanvasXpress Visualization of LHA summarized data 

**Dataset Preparation Notes**

* This RMD utilizes the SPEAK survey v1 (summarized) data spreadsheet  
* Names had to be cleaned to match up with the IDs in the data dictionary

```{r get_summary_data}
summary.raw <- read.xlsx(xlsxFile       = glue('{params$data_dir}/{params$summary_data_file}'),
                        sheet           = 'Survey Data', 
                        colNames        = FALSE,
                        skipEmptyCols   = FALSE,
                        na.strings      = c("NA", "Suppressed", ""),
                        fillMergedCells = TRUE)

# get good column names we can use
colnames(summary.raw) <- gsub('(%)', '', summary.raw[2, ], fixed = T) %>%
    trimws() %>%
    make.names(unique = T)

# hold onto "row" metadata, this is the communities/names
survey.col.meta <- summary.raw[1, , drop = F] %>% 
    t() %>% as.data.frame() %>%
    rownames_to_column("Name") %>%
    rename(Area = `1`)

# drop the old rows we used to make column names and index the metadata
summary.raw <- summary.raw[-(1:2), ]

# keep the first two columns when we sort out the part of the health data we want (LHA)
keep <- list("Domain" = 1, "Indicator" = 2)

# subset to the LHA columns
summary.lha <- summary.raw[, c(unlist(keep), 
                               which(survey.col.meta$Area == "Sub Regions"))]

# match the LHA IDs for use with map regions
id_lookup <- read.xlsx(xlsxFile        = glue('{params$data_dir}/{params$data_dictionary_file}'),
                       sheet           = 'CHSA_LHA_HSDA_HA_lookup', 
                       colNames        = TRUE,
                       na.strings      = c("NA", "Suppressed", ""),
                       fillMergedCells = TRUE) %>%
    select(LHA, LHA_Name) %>%
    mutate(LHA_Name = make.names(LHA_Name, unique = T))

# raw data
data    <- summary.lha[, -unlist(keep)] %>% t() %>% as.data.frame()
colnames(data) <- make.names(colnames(data), unique = T)
    
# column/sample (location) annotations
smp.ann <- summary.lha[, names(keep)]
rownames(smp.ann) <- colnames(data)

# row/variable (indication) annotations
var.ann <- data.frame("LHA_Name" = rownames(data), stringsAsFactors = F) %>%
    left_join(id_lookup, by = "LHA_Name")
rownames(var.ann) <- rownames(data)

#cleanup unused variables
# rm(summary.raw, survey.col.meta, id_lookup, keep, summary.lha)
```

**GEO Data Notes Preparation Notes**

* Map boundary data was obtained from:
https://www2.gov.bc.ca/gov/content/data/geographic-data-services/land-use/administrative-boundaries/health-boundaries

* Data was downsampled to 1% of original using http://mapshaper.org

```{r geo_data}
# lat/long geojson
boundaries.lha <- read_json(glue('{params$data_dir}/{params$lha_boundaries_file}'))
```

* Visualization Notes

This visualization uses CanvasXpress (www.canvasxpress.org)


```{r sample_data_for_debugging}
rownames(data) <- var.ann$LHA
rownames(var.ann) <- var.ann$LHA

sel <- c("X4", "X6", "X8")
data.sm <- data[, sel]
smp.sm  <- smp.ann[sel, ]
var.sm  <- var.ann
```


```{r lha_initial}
# chart tooltip events
lha.events <- JS("{
  'mousemove' : function(o, e, t) {
                    if (o) {
                        console.log(o);
                        t.showInfoSpan(e, '<b>' + o.z.LOCAL_HLTH_AREA_NAME[0] + 
                                          '</b> <i>(' + o.z.LOCAL_HLTH_AREA_CODE[0] + ')</i><br/>' +
                                          'HA: ' + o.z.HLTH_AUTHORITY_NAME[0] + 
                                          ' <i>(' + o.z.HLTH_AUTHORITY_CODE[0] + ')</i>');
                    };
}}")

# just show the map data (geojson)
canvasXpress(
    data       = FALSE,
    graphType  = "Map",
    title      = "Local Health Authorities (LHA)",
    showLegend = FALSE,
    topoJSON   = boundaries.lha,
    colorBy    = "HLTH_AUTHORITY_NAME",
    events     = lha.events
)

# with the added example data

# waiting on ID join, working with the CX author
# canvasXpress(
#     data          = data.sm,
#     varAnnot      = var.sm,
#     smpAnnot      = smp.sm,
#     graphType     = "Map",
#     mapPropertyId = "ID_1",
#     title         = "Local Health Authorities (LHA)",
#     showLegend    = FALSE,
#     topoJSON      = boundaries.lha,
#     colorBy       = "HLTH_AUTHORITY_NAME",
#     events        = lha.events
# )

```
